---
title: "Prepare long forms of power outage data"
author: "Shenyue Jia"
output: html_notebook
---


This notebook shows how to convert the original dataset purchased from poweroutage.us to long forms by county and city.

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(lubridate)
library(zoo)
library(foreign)
library(here)
library(bayestestR)
```

# Read original poweroutage.us CSV files

```{r}
rm(list = ls())
```

- Unzip file

```{r}
ca_data <- read_csv(unzip(here("data_private",
                               "POUS_Export_Raw_CityByUtility_20170101_20201231.zip"),
                          "POUS_Export_Raw_CityByUtility_20170101_20201231.csv")) %>%
    filter(StateName == "California")

head(ca_data)
```

# Fill missing values

Fill some missing values using the following rules:

-   "Los Angeles Department of Water & Power" = "Los Angeles"
-   "Pasadena Water and Power" = "Los Angeles"
-   "Riverside Public Utilities Department" = "Riverside"
-   "Sacramento Municipal Utility District" = "Sacramento"
-   "San Diego Gas & Electric" = "San Diego"

```{r}
ca_data$CountyName[which(ca_data$CountyName == "Unknown" &
                           ca_data$UtilityName == "Los Angeles Department of Water & Power")] <- "Los Angeles"

ca_data$CountyName[which(ca_data$CountyName == "Unknown" &
                           ca_data$UtilityName == "Pasadena Water and Power")] <- "Los Angeles"

ca_data$CountyName[which(ca_data$CountyName == "Unknown" &
                           ca_data$UtilityName == "Riverside Public Utilities Department")] <- "Riverside"

ca_data$CountyName[which(ca_data$CountyName == "Unknown" &
                           ca_data$UtilityName == "Sacramento Municipal Utility District")] <- "Sacramento"

ca_data$CountyName[which(ca_data$CountyName == "Unknown" &
                           ca_data$UtilityName == "San Diego Gas & Electric")] <- "San Diego"

ca_data$CountyName[which(ca_data$CountyName == "Unknown" &
                           ca_data$UtilityName == "Imperial Irrigation District")] <- "Imperial"

```

Check how many rows still have `CountyName` as `Unknown`.

```{r}
ca_data %>%
  subset(CountyName == "Unknown") %>%
  group_by(UtilityName) %>%
  summarize(count = n())
```

# Fix missing value of `CountyName` field with CDP data

Use census designated places (CDP) data to match `CityName`

```{r}
cdp <- read.csv(
  here(
    "data_private",
    "CDPs_County_Names.csv"
  )
)

cdp <- cdp %>%
  select(NAME10,COUNTY_NAME)


ca_data <- left_join(ca_data,cdp,by = c("CityName" = "NAME10"))

ca_data$CountyName[which(ca_data$CountyName == "Unknown")] <- 
  as.character(ca_data$COUNTY_NAME[which(ca_data$CountyName == "Unknown")])

ca_data <- ca_data %>%
  select(-COUNTY_NAME)

```

Check how many rows still have `CountyName` as `Unknown`. This time we were able to fix all missing value issues in `CountyName` field.

```{r}
ca_data %>%
  subset(CountyName == "Unknown") %>%
  group_by(UtilityName) %>%
  summarize(count = n())
```

# Obtain correct flags for start/end of power outage

Flag missing values, replace them with NA, increment their time ---- Note that missing values will have *identical* time stamps; however the missing indicator means that the *next* API ping (ten minutes later) did not return a valid JSON. In other words, the timestamp of the missing value indicator is actually valid. We will put the missing value in the next ten- minute time slot.

Some rows with missing data have the row returned a zero `CustomerOut` show up first. To correct this error, if a missing data problem is detected, assign both rows with the `CustomerOut` value (max among X and 0). Later in the script, assign the `CustomerOut` of the row marked as "missing" to 0

Manually add a row if the last row in the group has a `CustomerOut`\>0 This extra row will be manipulated later to add the correct missing flag, and be assigned to zero. Zero will be carried over in the next step

```{r}
ca_data <- ca_data %>%
  group_by(UtilityName, StateName, CountyName, CityName) %>%
  group_modify(~ .x %>% 
                 add_row(!!! map(.x, last)))

ca_data <- ca_data %>%
  group_by(UtilityName, StateName, CountyName, CityName, RecordedDateTime) %>%
  mutate(missing = 0:(n() - 1),
         CustomersOut = max(CustomersOut)) %>%
  ungroup() %>%
  mutate(
    CustomersOut = ifelse(missing == 1, 0, CustomersOut),
    datetime = ifelse(
      missing == 1,
      RecordedDateTime + lubridate::minutes(10),
      RecordedDateTime
    )
  ) %>% 
  mutate(datetime = lubridate::as_datetime(datetime))

```

-   Make ten-minute time intervals.

```{r}
ca_data <- ca_data %>%
  mutate(
    year = lubridate::year(datetime),
    month = lubridate::month(datetime),
    day =  lubridate::day(datetime),
    hour = lubridate::hour(datetime),
    minute = 10 * floor(lubridate::minute(datetime) / 10)
  ) %>%
  mutate(date = lubridate::ymd_hm(sprintf(
    "%s-%s-%s %s:%s",
    year,
    month,
    day,
    hour,
    minute
  ))) %>% 
  select(-datetime, -hour, -day, -month, -year, -missing)

```

-   Calculate total minutes

```{r}
ca_data <- ca_data %>% 
  group_by(UtilityName, StateName, CountyName, CityName) %>% 
  mutate(total_minutes = as.integer(difftime(lead(date), date))) %>%
  mutate(total_minutes = ifelse(is.na(total_minutes), 
                                as.integer(difftime(lubridate::ymd_hm("2020-01-01 00:00"), date)), 
                                total_minutes)
  )
```

# Generate long form of data

- Function to generate a long form of data

```{r}
return_county_longform <- function(df, county_name = county_name) {
  sub_df <- df %>% 
    filter(CountyName == county_name)
  
  df_utility_info <- expand(sub_df %>%
      select(UtilityName, StateName, CountyName, CityName),
      date = seq(
        from = as.POSIXct("2017-01-01 00:00", tz = "UTC"),
        to = as.POSIXct("2019-12-31 23:50", tz = "UTC"),
        by = "10 min"
      ))
  
  sub_df_sub <- sub_df %>% select(date,CustomersOut)
 
  res <- left_join(df_utility_info,
                       sub_df %>%
                       select(date,
                              UtilityName,
                              StateName,
                              CountyName,
                              CityName,
                              CustomersOut)
  ) %>% 
    arrange(UtilityName, StateName, CountyName, CityName, date)
  
  unique_combos <- res %>% 
    select(-date, -CustomersOut) %>% 
    distinct()
  CustomersOutLOCF <- vector("list", NROW(unique_combos))
  for (i in 1:NROW(unique_combos)) {
    CustomersOutLOCF[[i]] <- res %>% 
      filter(UtilityName == unique_combos$UtilityName[i],
             StateName == unique_combos$StateName[i],
             CountyName == unique_combos$CountyName[i],
             CityName == unique_combos$CityName[i]
      ) %>% 
      pull(CustomersOut) %>% 
      zoo::na.locf(na.rm = FALSE)
  }
  res$CustomersOut_NA <- unlist(CustomersOutLOCF)
  
  res <- res %>%
      mutate(CustomersOut_NA = ifelse(is.na(CustomersOut_NA),0,
                                      CustomersOut_NA))
  
  # res
}

```

- Generate long form for each county in California

```{r}
county_names <- as.character(levels(as.factor(ca_data$CountyName)))
n_county <- length(county_names)

for (i in 1:n_county){
  # i = 9
  long_form <- return_county_longform(ca_data, county_names[i])

  # Aggregate to county level
  long_form_county <- long_form %>%
      group_by(date) %>%
      summarise(CountyName = first(CountyName),
                CustomersOut = sum(CustomersOut_NA,na.rm = TRUE)) %>%
      arrange(date,CountyName,CustomersOut)

  print(paste("Writing",county_names[i],"Long Form"))
  
  # Save the aggregated version to CSV
  write_csv(long_form_county, file = here(
    "result",
    "long_form_missing_flg_fixed_county",
    paste0(county_names[i],"_long_form.csv")))
  
  # Save the city-level version as RDS format
  saveRDS(long_form, file = here("result",
                                 "long_form_missing_flg_fixed_city",
                                 paste0(county_names[i],"_long_form.RDS")))
  
  
}


```