---
title: "Inspect the city name missing data"
author: Shenyue Jia
output: html_notebook
---

This notebook aims to investigate the lack of power outage >= 60 min at the city level.

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)

```

```{r}
rm(list = ls())
```

## Investigate Amador County as an example

- Amador County has 41 county-level power outage events that are at least 60 min long from 2017 to 2020
- Amador County has 0 city (CDP)-level power outage events that are at least 60 min long from 2017 to 2020
- The lack of city-level power outage >= 60 min needs investigation

### Check missing data issue

- Read the original long-form of Amador county
```{r}
lf_amador_cdp <- readRDS(here("result","long_form_missing_flg_fixed_city",
                              "Amador_long_form.RDS"))
head(lf_amador_cdp)
```

- Check utility company names
```{r}
lf_amador_cdp %>%
    select(UtilityName, CityName) %>%
    distinct()
```

- Los Angeles Department of Water & Power and Silver Lake should not be included in the dataframe of `lf_amador_cdp`
- Temporarily fix
    - Remove all rows of this city from `lf_amador_cdp`

```{r}
lf_amador_cdp <- lf_amador_cdp %>%
    filter(!UtilityName %in% c("Los Angeles Department of Water & Power"))

lf_amador_cdp %>%
    select(UtilityName, CityName) %>%
    distinct()
```

- Check how many rows do not have total population

```{r}
# read in total housing unit number for CDPs
total_hh <- read.csv(here(
  "data_private",
  "housing-units-cdp-2020-cnty-names-california.csv"),stringsAsFactors = FALSE)
total_hh %>% DT::datatable(rownames = FALSE)
```

- Join the table and count the missing data rows
```{r}
long_form_hh <- lf_amador_cdp %>%
      left_join(total_hh %>%
                    select(total.housing.units,NAME10,COUNTY_NAME) %>%
                    filter(COUNTY_NAME == 'Amador'),
                by=c("CityName"="NAME10")) %>%
      select(!c("COUNTY_NAME")) %>%
      filter(!is.na(total.housing.units))

head(long_form_hh)

```

- How many do not have a matched total population number?
```{r}
long_form_hh %>%
    filter(is.na(total.housing.units)) %>%
    distinct(UtilityName,CityName)
```

- Turns out there is no missing data issue in Amador County

```{r}
# clean the workspace
rm(lf_amador_cdp,long_form_hh)
```

## Check the missing data issue for all countis in California

- Investigate if there are obviously wrong match between `UtilityName` and `CountyName` based on the service area of each utility company in this


```{r}
# get all the file names
long_form_fnames <- list.files(path = here("result",
                                           "long_form_missing_flg_fixed_city"),
                               pattern = "\\.RDS$")

n_long_form <- length(long_form_fnames)
```

- Get all unique combinations of `UtilityName`, `CountyName`, `CityName`
```{r}
for (f in 1:n_long_form){
  
  if(any(grepl("package:plyr", search()))) 
      detach("package:plyr",unload = TRUE) 
    else 
        message("plyr not loaded")

  
  # Read long form data of all CDP by county
  long_form <- readRDS(here(
    "result",
    "long_form_missing_flg_fixed_city",
    long_form_fnames[f]))
  
  curr_county <- substr(long_form_fnames[f],1,nchar(long_form_fnames[f])-14)
  
  print(paste('Working on',curr_county))
  
  # Add total housing unit number of each city for current county
  cdp_no_hh_cnt <- long_form %>%
      left_join(total_hh %>%
                    select(total.housing.units,NAME10,COUNTY_NAME) %>%
                    filter(COUNTY_NAME == curr_county),
                by=c("CityName"="NAME10","CountyName" = "COUNTY_NAME")) %>%
      # select(!c("COUNTY_NAME")) %>%
      group_by(UtilityName,CountyName,CityName) %>%
      summarise(no_cdp_hh = sum(is.na(total.housing.units)))
  
  rm(long_form)
  
  # Get names of CDP for current county
  if(f == 1){
      df_cdp_comb <- cdp_no_hh_cnt
  }else{
      df_cdp_comb <- bind_rows(df_cdp_comb,
                               cdp_no_hh_cnt)
  }
  
}  
```

- How many cities do not have a matched CDP total population?

```{r}
na_cdp_county <- df_cdp_comb %>%
    group_by(CountyName) %>%
    mutate(ncdp = n()) %>%
    ungroup() %>%
    filter(no_cdp_hh>0) %>%
    group_by(UtilityName, CountyName) %>%
    summarise(count = n(),
              pct_cnt = round(count/ncdp,2)*100)
```

- Save the names of CDP without CDP total population
```{r}
na_cdp_names <- df_cdp_comb %>%
    filter(no_cdp_hh>0) %>%
    select(!c("no_cdp_hh"))

write_csv(na_cdp_names, here("result","cdp_no_total_hh.csv"))
```