---
title: "Generate city-level power outage"
author: Shenyue Jia
output: html_notebook
---

This notebook generates a new set of power outage events using a different way to screen the events affecting less than 0.05% population.

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)

```

- Clean workspace
```{r}
rm(list = ls())
```


# Obtain housing unit data at CDP level

- Customers are defined as housing units in this work.

```{r}
if(any(grepl("package:plyr", search()))) detach("package:plyr",unload = TRUE) else message("plyr not loaded")

total_hh <- read.csv(here(
  "data_private",
  "housing-units-cdp-2020-cnty-names-california.csv"),stringsAsFactors = FALSE)
total_hh %>% DT::datatable(rownames = FALSE)
```

# Generate power outage events

- Obtain long form file names

```{r}
long_form_fnames <- list.files(path = here("result",
                                           "long_form_missing_flg_fixed_city"),
                               pattern = "\\.RDS$")

n_long_form <- length(long_form_fnames)
```

## Function to extract power outage

- Function to separate long form time series into power outage events
    - Detailed steps are described in the comments of code block

```{r}
generate_tile_plot_df <- function(long_form,customers_out){
  
  param_grid <- expand.grid(
    c_out = customers_out
  )

  for (i in 1:NROW(customers_out)){

    # If Customer out of power = 0, use a different condition to filter the subset data
    if (param_grid$c_out[i] == 0){
      long_form %>%
        filter(CustomersOut == param_grid$c_out[i]) -> curr_ts
      
      
    }else{
      long_form %>%
        filter(CustomersOut >= param_grid$c_out[i],
        ) -> curr_ts
      
    }
    

    # Obtain flags to separate the entire time series 
    # 1) if the time stamp is not consecutive (b/c the curr_ts is a subset from the long-form)
    # curr_ts$FlagCustomer <- c(0,diff(curr_ts$CustomersOut_Sum)) != 0
    # Note that this is different from the previous vesion
    # of script, which also judges if there is a change in customer out
    
    curr_ts$FlagTime <- c(10,diff(curr_ts$date)) > 10
    
    # Combine the two flags to get a general flag to separate the curr_ts into segments
    # Each segment contains the same number of customers out of power & the timestamps are consecutive
    curr_ts <- curr_ts %>%
      mutate(Flag = ifelse((FlagTime),TRUE,FALSE)) %>%
      select(-FlagTime)
    
    n_seg <- length(which(curr_ts$Flag))+1
    
    curr_ts$SegID <- NA
    curr_ts$SegID[which(curr_ts$Flag)] <- seq(1:n_seg)
    
    # Add back the first flag to the first row of curr_ts
    curr_ts$SegID[1] <- 0
    curr_ts$SegID = curr_ts$SegID+1
    curr_ts$SegID <- as.factor(curr_ts$SegID)
    
    # Number of segments (n_seg) = the number of consecutive rows with 
    # more than 0.5% total customers out of power (customers_out, varies)
    # by county
    # This value varies between each parameter combo
    
    curr_ts$NumSeg <- n_seg
    
    # Fill the segment ID
    curr_ts <- curr_ts %>% fill(SegID)
    
    # Record the maximum # of customer during the outage 
    # and the total minutes of duration (sum of segments * 10 min)
    # Duration in curr_ts = minutes of outage
    curr_ts <- curr_ts %>%
      group_by(SegID) %>%
      mutate(SegCustomerOut = max(CustomersOut,na.rm = TRUE),
             Duration = 10*n(),
             CustomerMinutes = sum(CustomersOut,na.rm = TRUE)*
               Duration)
    
    curr_ts <- curr_ts %>%
      group_by(SegID) %>%
      summarize(SegID = first(SegID),
                SegCustomerOut = first(SegCustomerOut),
                SegCustomerMinutes = first(CustomerMinutes),
                Duration = first(Duration),
                Start = first(date),
                End = last(date)+ms("9:59")) %>%
      mutate(CustomerThres = param_grid$c_out[i])
    
    # Calculate % of total housing units who have experienced power outage
    # Calculte % of time of the year when a power outage have occurred
    curr_ts <- curr_ts %>%
      mutate(PctHU = SegCustomerOut/curr_total_hh*100,
             PctTime = Duration/525600*100,
             TotalHU = curr_total_hh)
    
    # Drop unrealistic segmentations with % HU out of power > 100
    curr_ts <- curr_ts %>%
      filter(PctHU<100)
    
  }
  
  return(curr_ts)
}

```

- Loop over each county to convert long form data to power outage events

```{r}

for (f in 1:n_long_form){
  
  if(any(grepl("package:plyr", search()))) 
      detach("package:plyr",unload = TRUE) 
    else 
        message("plyr not loaded")

  
  # Read long form data of all CDP by county
  long_form <- readRDS(here(
    "result",
    "long_form_missing_flg_fixed_city",
    long_form_fnames[f]))
  
  curr_county <- substr(long_form_fnames[f],1,nchar(long_form_fnames[f])-14)
  
  # Add total housing unit number of each city for current county
  long_form_hh <- long_form %>%
      left_join(total_hh %>%
                    select(total_housing_unit,name_place,name_county) %>%
                    filter(name_county == curr_county),
                by=c("CityName"="name_place", "CountyName"="name_county")) %>%
      filter(!is.na(total_housing_unit))
 
  rm(long_form)
  
  # Get names of CDP for current county
  cdp_names <- unique(long_form_hh$CityName)
  
  if(length(cdp_names) != 0){
      
    for (cdp in 1:length(cdp_names)){
      
      curr_cdp <- cdp_names[cdp]
       # Get housing unit information from county level data `total_hh`
      curr_cdp_long_form <- long_form_hh[startsWith(long_form_hh$CityName,curr_cdp),]
      curr_total_hh <- curr_cdp_long_form$total_housing_unit[1]
      
      # # No limit on housing unit
      # customers_out <- 1
      
      # Apply limit on housing unit
      customers_out <- ifelse(round(curr_total_hh*0.005,0)<500,
                              500,
                              round(curr_total_hh*0.005,0))
      
      print(paste0("Working on ",curr_cdp,", ",curr_county," County"))
      
      curr_df <- generate_tile_plot_df(curr_cdp_long_form,customers_out)
  
      # Prepare saving to external file
      # Binding results for each county together
  
      col_names <- names(curr_df)
      curr_df <- bind_cols(curr_cdp,curr_county,curr_df)
      names(curr_df) <- c("CDP","County",col_names)
      if(cdp == 1){
        df_out <- curr_df
      }else{
        df_out <- bind_rows(df_out,curr_df)
      } 
  }
    
    # # Apply 0.5% cutoff and correct the cutoff value
    # df_out <- df_out %>%
    #     filter(PctHU > 0.05) %>%
    #     mutate(CustomerThres = round(TotalHU*0.05,0))
    
    df_out_sub <- df_out %>%
          filter(Duration >= 60)
      
    write_csv(df_out,
          here("result","power-outage-events-city",
               paste0(curr_county,"_cdp_pct_housing_unit_2017_2020.csv")))

    write_csv(df_out_sub,
          here("result","power-outage-events-city-at-least-60min",
               paste0(curr_county,"_cdp_pct_housing_unit_2017_2020.csv")))
      
  }
}                               

```


