---
title: "Generate a calendar heatmap for all power outages in California"
output: html_notebook
---

This notebook generates a calendar heat map for all county-level and city-level power outage events in California from 2017 to 2020

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(ggthemes)
library(hrbrthemes)
library(classInt)
library(stringr)
```

## Prepare data for plotting

- All county-level power outages
```{r}
df_cnty <- read_csv(here("result",
                         "ca_cnty_pct_housing_unit_at_least_60_min_2017_2020.csv"))
head(df_cnty)

```

- Use the selected power outage info above and the long-form data to generate a new version of long-form data with customers out of power

```{r}
county_names <- unique(df_cnty$County)

for (i in 1:length(county_names)){
    
    print(paste("Processing",county_names[i]))
    curr_lf <- read_csv(here("result","long_form_missing_flg_fixed_county",
                             paste0(county_names[i],"_long_form.csv")))
    
    # get a subset of power outage events
    curr_oop <- df_cnty %>%
        filter(County == county_names[i])
    
    # subset long form to keep the selected power outage events
    curr_lf_clean <- curr_lf %>%
        mutate(CustomersOut = ifelse(date >= curr_oop$Start & date <= curr_oop$End,
                                     CustomersOut, 0),
               TotalHU = first(curr_oop$TotalHU))
    
    rm(curr_lf)
    
    if(i == 1){
        lf_cnty_clean <- curr_lf_clean
    }else{
        lf_cnty_clean <- bind_rows(lf_cnty_clean,
                                   curr_lf_clean)
    }
    
}

```

- Generate `Year`,`Month`,`Day` for calendar heatmap

```{r}
lf_cnty_clean <- lf_cnty_clean %>%
    mutate(
    year = lubridate::year(date),
    month = lubridate::month(date),
    day =  lubridate::day(date),
    hour = lubridate::hour(date),
    minute = 10 * floor(lubridate::minute(date) / 10)
  ) %>%
  mutate(ymd = lubridate::ymd(sprintf(
    "%s-%s-%s",
    year,
    month,
    day
  )),
  md = format(as.Date(ymd), "%m-%d")) %>%
  select(-minute, -hour, -day, -month)

```

- Get a total number of housing units in California
```{r}
ca_tot_hu <- sum(unique(df_cnty$TotalHU))
```

- Aggregate data for calendar heatmap

```{r}

lf_cnty_clean_sum <- lf_cnty_clean %>%
  group_by(ymd) %>%
  summarise(CustomersOut = sum(CustomersOut,na.rm = TRUE),
            year = first(year),
            md = first(md),
            ymd = first(ymd)) %>%
  ungroup() %>%
    mutate(PctCustomerOut= CustomersOut/ca_tot_hu*100,
           year = as.numeric(format(ymd, format ="%Y")),
           md = format(ymd, format ="%m-%d"))

```

- Aggregate data by week for calendar heatmap
```{r}

lf_cnty_clean_sum_wk <- lf_cnty_clean %>%
    mutate(month = format(ymd,format = "%b"),
           week = format(ymd, format = "%W")) %>%
  group_by(week,year) %>%
  summarise(CustomersOut = sum(CustomersOut,na.rm = TRUE),
            month = first(month),
            ymd = first(ymd)) %>%
  ungroup() %>%
    mutate(PctCustomerOut= CustomersOut/ca_tot_hu*100)

```
- Remove week `00`
```{r}
lf_cnty_clean_sum_wk <- lf_cnty_clean_sum_wk %>%
    filter(!week %in% c('00','53') & ymd >= as.Date("2017-10-01"))
```

- Set the data of 2017 before the first non-zero value as `NA` as this dataset does not begin until Oct. 2017

```{r}
idx_nozero <- which(lf_cnty_clean_sum_wk$CustomersOut >0 )
lf_cnty_clean_sum_wk <- lf_cnty_clean_sum_wk[idx_nozero[1]:idx_nozero[length(idx_nozero)],]
```

## Create calendar heatmap plot

- Prepare a function to create beautiful breaks
```{r}
  caculate_class <- function(var,nclass){
    # library(plyr)
    class_brks <- classIntervals(var, n = nclass,intervalClosure = "left",
                                 style = "pretty")
    # class <- round_any(class_brks$brks,5)
    class <- class_brks$brks
    
    return(class)
  }
```

### Create calendar heatmap for weekly aggregation

- Prepare a vector for X-axis label
```{r}
df_mo_wk <- lf_cnty_clean_sum_wk %>%
    group_by(week) %>%
    summarise(month = first(month))

week_breaks <- df_mo_wk$week

week_labels <- rep("",nrow(df_mo_wk))
week_labels[seq(1:(nrow(df_mo_wk)/4))*4-2] = 
    df_mo_wk$month[seq(1:(nrow(df_mo_wk)/4))*4-2]
```

- Plot heatmap
```{r}
lf_cnty_clean_sum_wk %>%
  ggplot(aes(x=week, y=year, fill= PctCustomerOut)) + 
  geom_tile(color = "white",size = 0.1)+
  scale_fill_distiller(palette = "YlGnBu",trans = "reverse",name="% Customers\nlost power",
                       breaks = caculate_class(lf_cnty_clean_sum$PctCustomerOut,5),
                       labels = caculate_class(lf_cnty_clean_sum$PctCustomerOut,5)) +
  scale_x_discrete(breaks = week_breaks,
                   labels = week_labels) +

  scale_y_reverse()+
  theme_ipsum()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_line(color = "#cccccc",size = 0.1))+
  ggtitle("% Customers experiencing power outages in California each week") -> ca_heatmap

ca_heatmap
```

- Save results
```{r}
ggsave(
  here("plot","01_ca_power_outage_calendar_heatmap.png"),
  ca_heatmap,
  # device = cairo_pdf,
  width = 9,
  height = 6,
  scale = 1
)
```
