---
title: "Identify wildfire-induced events"
author: Shenyue Jia
output: html_notebook
---

## Introduction
- This notebook produce a dataset with a new flag field to identify if a power outage event is wildfire-induced.
- Wildfire-induced power outages
    - Directly caused by wildfire progression
    - Caused by Public Safety Power Shutoffs (PSPS)
    - Power outages caused by PSPS usually have a much larger geographical extent and higher number of customers affected

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(stringr)
```


- Clean workspace
```{r}
rm(list = ls())
```

## Get wildfire and PSPS events data
### Process data of PSPS at county level
```{r}
psps_cnty <- read_csv(here("data_private",
                           "psps-events-county-2017-2020.csv"))
head(psps_cnty)
```

- Format `psps_cnty` start/end date
```{r}
psps_cnty <- psps_cnty %>%
    mutate(`PSPS Start Date` = as.Date(`PSPS Start Date`, format = "%m/%d/%y"),
           `PSPS End Date` = as.Date(`PSPS End Date`, format = "%m/%d/%y"))
head(psps_cnty)
```

- Rename `psps_cnty` fields
```{r}
psps_cnty <- psps_cnty %>%
    rename("psps_start" = "PSPS Start Date",
           "psps_end" = "PSPS End Date")
names(psps_cnty)
```

### Process wildfire data at county level
```{r}
fire_cnty <- read_csv(here("data_private",
                           "ca-fire-2017-2022-place-name-county-name.csv"))
head(fire_cnty)
```

- Select fields and format dates
```{r}
fire_cnty <- fire_cnty %>%
    select(FIRE_NAME,ALARM_DATE,CONT_DATE,COUNTYFP,CNTY_FIPS,CNTY_NAME) %>%
    rename("fire_start" = "ALARM_DATE",
           "fire_end" = "CONT_DATE",
           "County" = "CNTY_NAME")

```

- Make sure there is no duplicated fire event in `fire_cnty`
```{r}
fire_cnty <- fire_cnty %>%
    distinct()
head(fire_cnty)
```



### Obtain county-level power outage events

```{r}
df_outage <- read_csv(here("result",
                           "ca_cnty_pct_housing_unit_at_least_60_min_2017_2020.csv"))
head(df_outage)
```

- Add date-only fields for `df_outage`

```{r}
df_outage <- df_outage %>%
    mutate(start_date = as.Date(Start),
           end_date = as.Date(End),
           event_id = seq(1:nrow(df_outage)))
head(df_outage)
```

### Prepare a new `dataframe` to include all wildfire-related events
```{r}
df_fire_event_cnty <- bind_rows(
    psps_cnty %>%
        select(County, psps_start, psps_end) %>%
        mutate(type = "PSPS") %>%
        rename(start_date = psps_start,
           end_date = psps_end),
    fire_cnty %>%
        select(County, fire_start, fire_end) %>%
        rename(start_date = fire_start,
               end_date = fire_end) %>%
        mutate(type = "Wildfire")
) %>%
    arrange(County)
head(df_fire_event_cnty)
```

## Identify county-level events using wildfire and PSPS info

- A function to judge of one power outage event is wildfire-induced or not
```{r}
# Function to check if a date is within a given period
is_date_within_period <- function(check_date, start_date, end_date) {
  check_date >= start_date & check_date <= end_date
}
```

- Loop to apply the function
```{r}
county_names <- unique(df_fire_event_cnty$County)

for (i in 1:length(county_names)){
    
    print(paste("Processing",county_names[i]))
    df_out <- df_outage %>%
        filter(County == county_names[i]) %>%
        arrange(start_date)
    
    df_fire <- df_fire_event_cnty %>%
        filter(County == county_names[i]) %>%
        arrange(start_date)
    
    for (j in 1:nrow(df_fire)){
        
        check_date <- df_fire$start_date[j]
        type <- df_fire$type[j]
        
        # Check if the specific date is within any of the periods
        result <- df_out %>%
          mutate(
            date_within_period = is_date_within_period(check_date, 
                                                       start_date, 
                                                       end_date),
          ) %>%
        filter(date_within_period == TRUE) %>%
            mutate(type = type)
        
        # bind rows for final results
        if (j == 1){
            result_bind <- result
        }else{
            result_bind <- bind_rows(
                result_bind,
                result
            )
        }
        
    }
    
    if (i == 1){
        df_fin <- result_bind
    }else{
        df_fin <- bind_rows(
            df_fin,
            result_bind
        )
    }
        
    
}

# remove some variables in the workspace
rm(df_out, result, result_bind)
```

- Apply table join to also flag power outages events that are not associated with wildfires or PSPS

```{r}
df_fin <- df_fin %>%
    select(event_id,date_within_period,type)

df_outage <- df_outage %>%
    left_join(df_fin, by = "event_id") %>%
    mutate(date_within_period = ifelse(is.na(date_within_period),
                                       "FALSE",date_within_period),
           type = ifelse(is.na(type),"Not Wildfire-induced",
                         type)) %>%
    select(-event_id, -date_within_period)
```

### Produce a county-level events count data for wildfire-induced power outages

- Remove the duplicated events again
```{r}
df_outage <- df_outage %>%
    distinct()
```

- Total number of events in each type
    - Wildfire
    - PSPS
    - Not wildfire-induced
```{r}
df_outage_cnty_cnt <- df_outage %>%
    group_by(County, type) %>%
    summarise(Count = n(),
              CustomerOut = sum(SegCustomerOut),
              CustomerMinutes = sum(SegCustomerMinutes),
              TotalDuration = sum(Duration),
              AvgDuration = mean(Duration),
              MedDuration = median(Duration),
              MaxDuration = max(Duration),
              MinDuration = min(Duration))
```


## Identify city-level events using city-level wildfire info

## Identify city-level events using county-level wildfire info
