---
title: "Identify county-level wildfire-induced events"
author: Shenyue Jia
output: html_notebook
---

## Introduction
- This notebook produce a dataset with a new flag field to identify if a power outage event at the CDP level is wildfire-induced.
- Wildfire-induced power outages
    - Directly caused by wildfire progression
    - Caused by Public Safety Power Shutoffs (PSPS)
    - Power outages caused by PSPS usually have a much larger geographical extent and higher number of customers affected

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(stringr)
```

- Clean workspace
```{r}
rm(list = ls())
```

## Obtain city-level wildfire-related events data

### Obtain wildfire events data
- Obtain cdp-level fire events
```{r}
fire_cdp <- read_csv(here("data_private",
                     "ca-fire-2017-2022-place-name-county-name.csv"))
head(fire_cdp)
```

```{r}
fire_cdp <- fire_cdp %>%
    select(CNTY_NAME,PLACE_FIPS, PLACE_NAME, PLACE_NAME_TYPE,
           ALARM_DATE,CONT_DATE) %>%
    rename("start_date" = "ALARM_DATE",
           "end_date" = "CONT_DATE",
           "county_name" = "CNTY_NAME",
           "place_fips" = "PLACE_FIPS",
           "place_name" = "PLACE_NAME",
           "place_name_type" = "PLACE_NAME_TYPE") %>%
    mutate(place_fips = paste0("0",place_fips),
           type = "Wildfire") %>%
    arrange(county_name,place_name)

head(fire_cdp)

```

- Make sure there is no duplicated fire event in `fire_cdp`
```{r}
fire_cdp <- fire_cdp %>%
    distinct()
head(fire_cdp)
```

### Obtain PSPS events data at county level

- The dataframe structure will be **different** than the `fire_cdp`
    - No `place_fips`, `place_name`, `place_name_type`
    
```{r}
psps_cnty <- read_csv(here("data_private","psps-events-county-2017-2020.csv"))
head(psps_cnty)
```

- Format `psps_cnty` start/end date
```{r}
psps_cnty <- psps_cnty %>%
    mutate(`PSPS Start Date` = as.Date(`PSPS Start Date`, format = "%m/%d/%y"),
           `PSPS End Date` = as.Date(`PSPS End Date`, format = "%m/%d/%y"))
head(psps_cnty)
```

- Rename fields and re-organized fields in `psps_cnty`
```{r}
psps_cnty <- psps_cnty %>%
    rename("start_date" = "PSPS Start Date",
           "end_date" = "PSPS End Date",
           "county_name" = "County") %>%
    mutate(type = "PSPS") %>%
    select(county_name,start_date,end_date,type) %>%
    arrange(county_name)
names(psps_cnty)
head(psps_cnty)
```

- Make sure there is no duplicated fire event in `psps_cnty`
```{r}
psps_cnty <- psps_cnty %>%
    distinct()
head(psps_cnty)
```

## Identify city-level events using city and county-level wildfire-related events

- A function to judge of one power outage event is wildfire-induced or not
```{r}
# Function to check if a date is within a given period
is_date_within_period <- function(check_date, start_date, end_date) {
  check_date >= start_date & check_date <= end_date
}
```

- Loop to apply the function and flag if an event is wildfire-induced or not
- This time, we need to apply different treatment for PSPS and wildfire events as PSPS data is only available at **county** level

- A function to judge if an event is wildfire-induced

```{r}
wildfire_induced_events <- function(curr_county,
                                    curr_df_outage, curr_df_fire){
    for (j in 1:nrow(curr_df_fire)){
                
                check_date <- curr_df_fire$start_date[j]
                type <- curr_df_fire$type[j]
                
                # Check if the specific date is within any of the periods
                result <- curr_df_outage %>%
                  mutate(
                    date_within_period = is_date_within_period(check_date, 
                                                               Start, 
                                                               End),
                  ) %>%
                filter(date_within_period == TRUE) %>%
                    mutate(type = type)
                
                # bind rows for final results
                if (j == 1){
                    result_bind <- result
                }else{
                    result_bind <- bind_rows(
                        result_bind,
                        result
                    )
                }
                
    }
    
    return(result_bind)
    
}
```

- Obtain cdp-level data for each county
```{r}
cdp_file_names <- list.files(here("result",
                                  "power-outage-events-city-at-least-60min"),
                             pattern = "\\.csv$", ignore.case = TRUE)
county_names <- substr(cdp_file_names,1, 
                       (nchar(cdp_file_names)-nchar("_cdp_pct_housing_unit_2017_2020.csv")))
```


### Check if any of the cdp-level event is wildfire-induced


- Apply judge function for wildfire events

```{r}
for (i in 1:length(county_names)){
    
    print(paste("Processing",county_names[i]))
    
    # Obtain data for city-level power outage
    # City-level power outage data are organized by counties

    df_outage <- read_csv(here("result",
                               "power-outage-events-city-at-least-60min",
                               paste(county_names[i],
                                     "cdp_pct_housing_unit_2017_2020.csv",
                                     sep = "_")))
    
    if(nrow(df_outage)>0){
        
        df_outage <- df_outage %>%
            arrange(Start)
        
        df_fire <- fire_cdp %>%
            filter(county_name == county_names[i]) %>%
            arrange(start_date, county_name)
        
        if(nrow(df_fire)>0){
            curr_result <- wildfire_induced_events(county_name[i],
                                               df_outage,
                                               df_fire)            
        }
        
    # bind results for counties together
    if(i==1){
        out_df_wildfire <- curr_result
    }else{
        out_df_wildfire <- bind_rows(out_df_wildfire,
                            curr_result)
    }    
        

        
        
        
    }
    
    
}

```

### Check if any of the cdp-level event is PSPS-induced

- Apply the function to determine if a power outage event is a wildfire-induced one
```{r}
for (i in 1:length(county_names)){
    
    print(paste("Processing",county_names[i]))
    
    # Obtain data for city-level power outage
    # City-level power outage data are organized by counties

    df_outage <- read_csv(here("result",
                               "power-outage-events-city-at-least-60min",
                               paste(county_names[i],
                                     "cdp_pct_housing_unit_2017_2020.csv",
                                     sep = "_")))
    
    if(nrow(df_outage)>0){
        
        df_outage <- df_outage %>%
            arrange(Start)
        
        df_fire <- psps_cnty %>%
            filter(county_name == county_names[i]) %>%
            arrange(start_date, county_name)
        
        if(nrow(df_fire)>0){
            curr_result <- wildfire_induced_events(county_name[i],
                                               df_outage,
                                               df_fire)            
        }
        
    # bind results for counties together
    if(i==1){
        out_df_psps <- curr_result
    }else{
        out_df_psps <- bind_rows(out_df_psps,
                            curr_result)
    }    
        

        
        
        
    }
    
    
}

```

## Note on results
- Both `out_df_wildfire` and `out_df_psps` are empty, indicating no valid results are available and can be marked as PSPS or wildfire-induced.
- This needs further investigation, including
    - Checking the long form of time series
    - Checking the threshold and other parameters applied to identify city-level events

