---
title: "Identify county-level wildfire-induced events"
author: Shenyue Jia
output: html_notebook
---

## Introduction
- This notebook produce a dataset with a new flag field to identify if a power outage event at the CDP levelis wildfire-induced.
- Wildfire-induced power outages
    - Directly caused by wildfire progression
    - Caused by Public Safety Power Shutoffs (PSPS)
    - Power outages caused by PSPS usually have a much larger geographical extent and higher number of customers affected

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(stringr)
```

- Clean workspace
```{r}
rm(list = ls())
```

## Obtain city-level wildfire-related events data

### Obtain wildfire events data
- Obtain cdp-level fire events
```{r}
fire_cdp <- read_csv(here("data_private",
                     "ca-fire-2017-2022-place-name-county-name.csv"))
head(fire_cdp)
```

```{r}
fire_cdp <- fire_cdp %>%
    select(CNTY_NAME,PLACE_FIPS, PLACE_NAME, PLACE_NAME_TYPE,
           ALARM_DATE,CONT_DATE) %>%
    rename("start_date" = "ALARM_DATE",
           "end_date" = "CONT_DATE",
           "county_name" = "CNTY_NAME",
           "place_fips" = "PLACE_FIPS",
           "place_name" = "PLACE_NAME",
           "place_name_type" = "PLACE_NAME_TYPE") %>%
    mutate(place_fips = paste0("0",place_fips),
           type = "Wildfire") %>%
    arrange(county_name,place_name)

head(fire_cdp)

```

- Make sure there is no duplicated fire event in `fire_cdp`
```{r}
fire_cdp <- fire_cdp %>%
    distinct()
head(fire_cdp)
```

### Obtain PSPS events data at county level

- The dataframe structure will be **different** than the `fire_cdp`
    - No `place_fips`, `place_name`, `place_name_type`
    
```{r}
psps_cnty <- read_csv(here("data_private","psps-events-county-2017-2020.csv"))
head(psps_cnty)
```

- Format `psps_cnty` start/end date
```{r}
psps_cnty <- psps_cnty %>%
    mutate(`PSPS Start Date` = as.Date(`PSPS Start Date`, format = "%m/%d/%y"),
           `PSPS End Date` = as.Date(`PSPS End Date`, format = "%m/%d/%y"))
head(psps_cnty)
```

- Rename fields and re-organized fields in `psps_cnty`
```{r}
psps_cnty <- psps_cnty %>%
    rename("start_date" = "PSPS Start Date",
           "end_date" = "PSPS End Date",
           "county_name" = "County") %>%
    mutate(type = "PSPS") %>%
    select(county_name,start_date,end_date,type) %>%
    arrange(county_name)
names(psps_cnty)
head(psps_cnty)
```

- Make sure there is no duplicated fire event in `psps_cnty`
```{r}
psps_cnty <- psps_cnty %>%
    distinct()
head(psps_cnty)
```

## Identify city-level events using city and county-level wildfire-related events

- A function to judge of one power outage event is wildfire-induced or not
```{r}
# Function to check if a date is within a given period
is_date_within_period <- function(check_date, start_date, end_date) {
  check_date >= start_date & check_date <= end_date
}
```

- Loop to apply the function and flag if an event is wildfire-induced or not
- This time, we need to apply different treatment for PSPS and wildfire events as PSPS data is only available at **county** level

### Check if any of the cdp-level event is wildfire-induced
```{r}
cdp_file_names <- list.files(here("result",
                                  "power-outage-events-city-at-least-60min"),
                             pattern = "\\.csv$", ignore.case = TRUE)
county_names <- substr(cdp_file_names,1, 
                       (nchar(cdp_file_names)-nchar("_cdp_pct_housing_unit_2017_2020.csv")))

for (i in 1:length(county_names)){
    
    print(paste("Processing",county_names[i]))
    
    # Obtain data for city-level power outage
    # City-level power outage data are organized by counties
    
    
    df_outage <- read_csv(here("result",
                               "power-outage-events-city-at-least-60min",
                               paste(county_names[i],
                                     "cdp_pct_housing_unit_2017_2020.csv",
                                     sep = "_")))
    
    if(nrow(df_outage) > 0){
            
        df_outage <- df_outage %>% arrange(Start)
    
    
        df_fire <- fire_cdp %>%
                filter(county_name == county_names[i]) %>%
                arrange(start_date, place_name)
        
        if(nrow(df_fin) > 0){
           for (j in 1:nrow(df_fire)){
                
                check_date <- df_fire$start_date[j]
                type <- df_fire$type[j]
                
                # Check if the specific date is within any of the periods
                result <- df_outage %>%
                  mutate(
                    date_within_period = is_date_within_period(check_date, 
                                                               Start, 
                                                               End),
                  ) %>%
                filter(date_within_period == TRUE) %>%
                    mutate(type = type)
                
                # bind rows for final results
                if (j == 1){
                    result_bind <- result
                }else{
                    result_bind <- bind_rows(
                        result_bind,
                        result
                    )
                }
                
            }
            
           if (i == 1){
                df_fin <- result_bind
            }else{
                df_fin <- bind_rows(
                    df_fin,
                    result_bind
                )
            }    
        }
            

        
    }    
        

        
    
}

# remove some variables in the workspace
rm(df_out, result, result_bind)
```

### Check if any of the cdp-level event is PSPS-induced
```{r}
cdp_file_names <- list.files(here("result",
                                  "power-outage-events-city-at-least-60min"),
                             pattern = "\\.csv$", ignore.case = TRUE)
county_names <- substr(cdp_file_names,1, 
                       (nchar(cdp_file_names)-nchar("_cdp_pct_housing_unit_2017_2020.csv")))

for (i in 1:length(county_names)){
    
    print(paste("Processing",county_names[i]))
    
    # Obtain data for city-level power outage
    # City-level power outage data are organized by counties
    
    
    df_outage <- read_csv(here("result",
                               "power-outage-events-city-at-least-60min",
                               paste(county_names[i],
                                     "cdp_pct_housing_unit_2017_2020.csv",
                                     sep = "_")))
    
    if(nrow(df_outage) > 0){
            
        df_outage <- df_outage %>% arrange(Start)
    
    
        df_fire <- psps_cnty %>%
                filter(county_name == county_names[i]) %>%
                arrange(start_date, county_name)
        
        if(nrow(df_fin) > 0){
           for (j in 1:nrow(df_fire)){
                
                check_date <- df_fire$start_date[j]
                type <- df_fire$type[j]
                
                # Check if the specific date is within any of the periods
                result <- df_outage %>%
                  mutate(
                    date_within_period = is_date_within_period(check_date, 
                                                               Start, 
                                                               End),
                  ) %>%
                filter(date_within_period == TRUE) %>%
                    mutate(type = type)
                
                # bind rows for final results
                if (j == 1){
                    result_bind <- result
                }else{
                    result_bind <- bind_rows(
                        result_bind,
                        result
                    )
                }
                
            }
            
           if (i == 1){
                df_fin <- result_bind
            }else{
                df_fin <- bind_rows(
                    df_fin,
                    result_bind
                )
            }    
        }
            

        
    }    
        

        
    
}

# remove some variables in the workspace
rm(df_out, result, result_bind)
```
