---
title: "Prepare aggregated data for power outage at county-level"
author: Shenyue Jia
output: html_notebook
---

This notebook prepared aggregated data from county-level power outage events. 

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(ggthemes)
library(hrbrthemes)
library(classInt)
library(stringr)
```

- Clean workspace
```{r}
rm(list = ls())
```


## Read county-level power outage data with type flag

- Total number of events in each type
    - Wildfire
    - PSPS
    - Not wildfire-induced

```{r}
df_outage <- read_csv(here("result","county-outage-events-by-type.csv"))
```

## Prepare aggregated results

### `Count` and `type`
- Total count of events by county
```{r}
df_outage_cnty_cnt <- df_outage %>%
    group_by(County) %>%
    summarise(Count = n(),
              TotalHU = first(TotalHU),
              CustomerOut = sum(SegCustomerOut),
              CustomerMinutes = sum(SegCustomerMinutes),
              MedCustomerMinutes = median(SegCustomerMinutes),
              MinCustomerMinutes = min(SegCustomerMinutes),
              MaxCustomerMinutes = max(SegCustomerMinutes),
              AvgCustomerMinutes = mean(SegCustomerMinutes),
              MedCustomerOut = median(SegCustomerOut),
              MinCustomerOut = min(SegCustomerOut),
              MaxCustomerOut = max(SegCustomerOut),
              AvgCustomerOut = mean(SegCustomerOut),
              TotalDuration = sum(Duration),
              AvgDuration = mean(Duration),
              MedDuration = median(Duration),
              MaxDuration = max(Duration),
              MinDuration = min(Duration),
              PctCustomerOut = CustomerOut/(4*TotalHU),
              PctDuration = round(TotalDuration/(525600*4),4)*100)
```

```{r}
write_csv(df_outage_cnty_cnt, here("result","county-aggregation-results",
                                   "county-poweroutage-count-2017-2020.csv"))
```

- Total count of events by county and type
```{r}
df_outage_cnty_type_cnt <- df_outage %>%
    group_by(County, type) %>%
    summarise(Count = n(),
              TotalHU = first(TotalHU),
              CustomerOut = sum(SegCustomerOut),
              CustomerMinutes = sum(SegCustomerMinutes),
              MedCustomerMinutes = median(SegCustomerMinutes),
              MinCustomerMinutes = min(SegCustomerMinutes),
              MaxCustomerMinutes = max(SegCustomerMinutes),
              AvgCustomerMinutes = mean(SegCustomerMinutes),
              MedCustomerOut = median(SegCustomerOut),
              MinCustomerOut = min(SegCustomerOut),
              MaxCustomerOut = max(SegCustomerOut),
              AvgCustomerOut = mean(SegCustomerOut),
              TotalDuration = sum(Duration),
              AvgDuration = mean(Duration),
              MedDuration = median(Duration),
              MaxDuration = max(Duration),
              MinDuration = min(Duration),
              PctCustomerOut = CustomerOut/(4*TotalHU),
              PctDuration = round(TotalDuration/(525600*4),4)*100)
```

- Calculate the correct `pct_customer` out of power using 3*Total Housing Unit of each county

```{r}
df_hu <- df_outage %>%
    group_by(County) %>%
    summarise(TotalHU = first(TotalHU))
head(df_hu)
```

```{r}
ca_total_hu <- sum(df_hu$TotalHU)
```


```{r}
write_csv(df_outage_cnty_type_cnt, here("result","county-aggregation-results",
                                   "county-poweroutage-count-type-2017-2020.csv"))
```

- Percent count of each `type` of power outages
```{r}
df_cnt_pct_type_sum <- df_outage %>%
    group_by(County) %>%
    summarise(total_count = n())

df_cnt_pct_type <- df_outage %>%
    group_by(County, type) %>%
    summarise(count = n()) %>%
    left_join(df_cnt_pct_type_sum, by = c("County")) %>%
    mutate(pct_total = round(count/total_count,4)*100) %>%
    ungroup()

df_cnt_pct_type <- df_cnt_pct_type %>%
    complete(County, type) %>%
    fill(total_count) %>%
    mutate(count = ifelse(is.na(count),0,count),
           # total_count = ifelse(is.na(total_count),total_count,total_count),
           pct_total = ifelse(is.na(pct_total),0,pct_total))
```

- Produce a new dataframe with sum of wildfire-induced events
```{r}
df_cnt_pct_wf_induced <- df_cnt_pct_type %>%
    filter(type %in% c("Wildfire","PSPS")) %>%
    group_by(County) %>%
    summarise(type = "Wildfire-Induced",
              count = sum(count),
              total_count = first(total_count),
              pct_total = sum(pct_total))
```

- Attach it to `df_cnt_pct_type`
```{r}
df_cnt_pct_type <- bind_rows(df_cnt_pct_type,
                             df_cnt_pct_wf_induced)
```

```{r}
write_csv(df_cnt_pct_type, here("result","county-aggregation-results",
                                "county-poweroutage-count-type-pct-2017-2020.csv"))

write_csv(df_cnt_pct_type %>%
              filter(type ==  "PSPS"), 
          here("result","county-aggregation-results",
                                "county-poweroutage-count-type-pct-psps-2017-2020.csv"))

write_csv(df_cnt_pct_type %>%
              filter(type == "Wildfire"), 
          here("result","county-aggregation-results",
                                "county-poweroutage-count-type-pct-wildfire-only-2017-2020.csv"))

write_csv(df_cnt_pct_type %>%
              filter(type == "Not Wildfire-induced"), 
          here("result","county-aggregation-results",
                                "county-poweroutage-count-type-pct-not-wildfire-2017-2020.csv"))

write_csv(df_cnt_pct_type %>%
              filter(type == "Wildfire-Induced"), 
          here("result","county-aggregation-results",
                                "county-poweroutage-count-type-pct-wildfire-induced-2017-2020.csv"))
```


- Total count of events: `PSPS`
```{r}
df_outage_psps <- df_outage_cnty_type_cnt %>%
    filter(type %in% c("PSPS"))
write_csv(df_outage_psps, here("result","county-aggregation-results",
                                   "county-poweroutage-count-psps-2017-2020.csv"))
```

- Total count of events: `Wildfire`
```{r}
df_outage_wildfire <- df_outage_cnty_type_cnt %>%
    filter(type %in% c("Wildfire"))
write_csv(df_outage_wildfire, here("result","county-aggregation-results",
                                   "county-poweroutage-count-wildfire-2017-2020.csv"))
```

- Total count of events: `Not Wildfire-induced`
```{r}
df_outage_notwildfire <- df_outage_cnty_type_cnt %>%
    filter(type %in% c("Not Wildfire-induced"))
write_csv(df_outage_notwildfire, here("result","county-aggregation-results",
                                   "county-poweroutage-count-not-wildfire-2017-2020.csv"))
```

- Total count of events by months
```{r}
# add month field
df_outage <- df_outage %>%
    mutate(month = format(Start, "%b"),
           month_num = format(Start, "%m"))
head(df_outage)
```

- Aggregate `df_outage` by `month` and `County`
```{r}
df_outage_mon_agg <- df_outage %>%
    group_by(month, County) %>%
    summarise(Count = n(),
              TotalHU = first(TotalHU),
              CustomerOut = sum(SegCustomerOut),
              CustomerMinutes = sum(SegCustomerMinutes),
              MedCustomerMinutes = median(SegCustomerMinutes),
              MinCustomerMinutes = min(SegCustomerMinutes),
              MaxCustomerMinutes = max(SegCustomerMinutes),
              AvgCustomerMinutes = mean(SegCustomerMinutes),
              MedCustomerOut = median(SegCustomerOut),
              MinCustomerOut = min(SegCustomerOut),
              MaxCustomerOut = max(SegCustomerOut),
              AvgCustomerOut = mean(SegCustomerOut),
              TotalDuration = sum(Duration),
              MaxDuration = max(Duration),
              MinDuration = min(Duration),
              AvgDuration = mean(Duration),
              MedDuration = median(Duration),
              PctCustomerOut = CustomerOut/(4*TotalHU),
              PctDuration = round(TotalDuration/(525600*4),4)*100)
```

```{r}
write_csv(df_outage_mon_agg, here("result","county-aggregation-results",
                                  "county-poweroutage-count-bymonth-2017-2020.csv"))
```

## Investigate events by `Duration`
### Define cut-off values by `Duration`

- Less than 8 hours
- Less than 24 hours
- Less than 72 hours

```{r}
df_outage <- df_outage %>%
    mutate(duration_cate = 
        case_when(Duration <= 8*60 ~ '<= 8 hours'
                  ,Duration >8*60 & Duration <= 24*60 ~ '9-24 hours'
                  ,Duration > 24*60 & Duration <= 72*60 ~ '25-72 hours'
                  ,Duration >= 72*60 ~ '72+ hours'
                  ,TRUE ~ 'Other'
                  )
        
    )

```

### Aggregate by `month` and `duration_cate`

- Aggregation over months
```{r}
df_outage_mon_dur_agg <- df_outage %>%
    group_by(month, month_num, duration_cate) %>%
    summarise(Count = n(),
              CustomerOut = sum(SegCustomerOut),
              CustomerMinutes = sum(SegCustomerMinutes),
              MedCustomerMinutes = median(SegCustomerMinutes),
              MinCustomerMinutes = min(SegCustomerMinutes),
              MaxCustomerMinutes = max(SegCustomerMinutes),
              AvgCustomerMinutes = mean(SegCustomerMinutes),
              MedCustomerOut = median(SegCustomerOut),
              MinCustomerOut = min(SegCustomerOut),
              MaxCustomerOut = max(SegCustomerOut),
              AvgCustomerOut = mean(SegCustomerOut),
              TotalDuration = sum(Duration),
              MaxDuration = max(Duration),
              MinDuration = min(Duration),
              AvgDuration = mean(Duration),
              MedDuration = median(Duration),
              TotalDuration) %>%
    arrange(month_num)
```

### Contribution of each type by month

- Calculate the total power outage events by month
```{r}
df_tot_outage <- df_outage %>%
    group_by(duration_cate) %>%
    summarise(total_count = n())
```

- Join `df_tot_outage` with `df_outage`
```{r}
df_outage_mon_dur_agg <- df_outage_mon_dur_agg %>%
    left_join(df_tot_outage, by = "duration_cate") %>%
    mutate(event_pct_total = round(Count/total_count,4)*100)
head(df_outage_mon_dur_agg)
```

- Save to external file
```{r}
write_csv(df_outage_mon_dur_agg, here("result","county-aggregation-results",
                                      "state-poweroutage-count-bymonth-2017-2020.csv"))
```

## Investigate `type` (wildfire-induced or not)
### Count of each type by `month` by `duration_cate`

```{r}
df_outag_mon_duration_type <- df_outage %>%
    group_by(month_num,month,duration_cate, type) %>%
    summarise(Count = n(),
              TotalHU = first(TotalHU),
              CustomerOut = sum(SegCustomerOut),
              CustomerMinutes = sum(SegCustomerMinutes),
              MedCustomerMinutes = median(SegCustomerMinutes),
              MinCustomerMinutes = min(SegCustomerMinutes),
              MaxCustomerMinutes = max(SegCustomerMinutes),
              AvgCustomerMinutes = mean(SegCustomerMinutes),
              MedCustomerOut = median(SegCustomerOut),
              MinCustomerOut = min(SegCustomerOut),
              MaxCustomerOut = max(SegCustomerOut),
              AvgCustomerOut = mean(SegCustomerOut),
              TotalDuration = sum(Duration),
              MaxDuration = max(Duration),
              MinDuration = min(Duration),
              AvgDuration = mean(Duration),
              MedDuration = median(Duration),
              PctCustomerOut = CustomerOut/(4*TotalHU),
              PctDuration = round(TotalDuration/(525600*4),4)*100) %>%
    arrange(month_num)
```

```{r}
write_csv(df_outag_mon_duration_type,
          here("result","county-aggregation-results",
               "state-poweroutage-count-bymonth-type-2017-2020.csv"))
```
