---
title: "Map county-level power outage results"
author: Shenyue Jia
output: html_notebook
---

This notebook maps out a few county-level aggregation results.

```{r setup, warning=FALSE, , message=FALSE}
library(rstudioapi)
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble
library(magrittr) # pipes
library(lintr) # code linting
library(sf) # spatial data handling
library(viridis) # viridis color scale
library(cowplot) # stack ggplots
library(rmarkdown)
library(here)
library(hrbrthemes)
library(ggExtra)
library(rgeoda)
```

- Clean up workspace
```{r}
rm(list = ls())
```

## Prepare spatial data for mapping

- Read Shapefiles of counties in California

```{r}
map_ca_county<- st_read(here("data_private",
                             "ca_county_census_empower_UDSMapper_combined.shp"))

names(map_ca_county)
head(map_ca_county)
```

- Drop some unused fields
```{r}
map_ca_county <- map_ca_county %>%
    select(STATEFP,COUNTYFP,GEOID,NAME)
head(map_ca_county)
```

## Define a map theme
```{r}
theme_map <- function(...) {
  theme_ipsum()+
  theme(
    # text = element_text(family = default_font_family,
    #                     color = default_font_color),
    # remove all axes
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    # add a subtle grid
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # background colors
    plot.background = element_rect(
      fill = NULL,
                                   color = NA),
    panel.background = element_rect(
      fill = NULL,
                                    color = NA),
    legend.background = element_rect(
      fill = NULL,
                                     color = NA),
    # borders and margins
    plot.margin = unit(c(0,0,0,0), "cm"),
    panel.border = element_blank(),
    panel.spacing = unit(c(0,0,0,0), "cm"),
    # titles
    legend.title = element_text(size = 11),
    legend.text = element_text(
      # color = default_font_color,
      size = 9, hjust = 0),
    plot.title = element_text(
      # color = default_font_color,
    size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 10, hjust = 0.5,
                                 # color = default_font_color,
                                 margin = margin(b = -0.1,
                                                 t = -0.1,
                                                 l = 2,
                                                 unit = "cm"),
                                 debug = F),
    # captions
    plot.caption = element_text(size = 7,
                                hjust = .5,
                                margin = margin(t = 0.2,
                                                b = 0,
                                                unit = "cm"),
                                color = "#939184"),
    ...
  )
}
```

- Prepare a function to create beautiful breaks
```{r}
  caculate_class <- function(var,nclass,style){
    # library(plyr)
    class_brks <- classIntervals(var, n = nclass,intervalClosure = "left",
                                 style = style)
    # class <- round_any(class_brks$brks,5)
    class <- class_brks$brks
    
    return(class)
  }
```

## Map count of power outage events by type at county-level

- Read aggregated county results
```{r}
df_metric <- read_csv(here("result","county-aggregation-results",
                           "county-poweroutage-count-2017-2020.csv"))
head(df_metric)
```

- Join metrics to `map_ca_county`

```{r}
map_df_metric <- map_ca_county %>%
    left_join(df_metric, by = c("NAME" = "County"))
head(map_df_metric)
```

- Save to a Shapefile
```{r}
st_write(df_metric, here("result","geography","county-outage-bytype.shp"))
```

- Create a map with `map_df_metric`

```{r}
map_count <- map_df_metric %>%
    ggplot()+
    geom_sf(
        aes(fill = Count),
        color = "white", size = 0.05
    )+
    scale_fill_distiller(palette = "YlGnBu",trans = "reverse",name="Count",
                       breaks = caculate_class(map_df_metric$Count,4,"jenks"),
                       labels = caculate_class(map_df_metric$Count,4,"jenks"))+
    # geom_sf(
    #     fill = "transparent",
    #     color = "white", size = 0.3
    # )+
    labs(x = NULL,
         y = NULL,
         title = "a) County-level power outages (2017-2020)")+
    theme_map()
map_count
```

```{r}
map_cust_oop <- map_df_metric %>%
    ggplot()+
    geom_sf(
        aes(fill = CustomerOut),
        color = "white", size = 0.05
    )+
    scale_fill_distiller(palette = "YlGnBu",trans = "reverse",name="Customers\nOut of Power",
                       breaks = caculate_class(map_df_metric$CustomerOut,4,"pretty"),
                       labels = format(caculate_class(map_df_metric$CustomerOut,4,"pretty"),
                                       big.mark = ","))+
    # geom_sf(
    #     fill = "transparent",
    #     color = "white", size = 0.3
    # )+
    labs(x = NULL,
         y = NULL,
         title = "b) Accumulated customers out of power (2017-2020)")+
    theme_map()
map_cust_oop
```
- Bivariate map