---
title: "Visualize emPOWER data"
author: Shenyue Jia
output: html_notebook
---

This notebook creates several visualizations for emPOWER data from 2017 to 2022

```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(ggthemes)
library(hrbrthemes)
library(classInt)
library(stringr)
```

- Clean workspace
```{r}
rm(list = ls())
```

## Read emPOWER data

```{r}
df_medicare <- read_csv(here("data_private","emPOWER",
                             "medicare_bene_long_form_2017_2022.csv")) %>%
    mutate(month_year = my(paste(month,year,sep = "-")))

df_dme <- read_csv(here("data_private","emPOWER",
                   "dme_long_form_2017_2022.csv")) %>%
    mutate(month_year = my(paste(month,year,sep = "-")))
```

## Generate calendar heatmap

- Prepare a function to create beautiful breaks
```{r}
  caculate_class <- function(var,nclass){
    # library(plyr)
    class_brks <- classIntervals(var, n = nclass,intervalClosure = "left",
                                 style = "pretty")
    # class <- round_any(class_brks$brks,5)
    class <- class_brks$brks
    
    return(class)
  }
```


- Aggregate data to California
```{r}
df_medicare_agg <- df_medicare %>%
    group_by(month_year) %>%
    summarise(value = sum(value),
           month = format(month_year,"%b"),
           year = as.numeric(format(month_year,"%Y")),
           month_num = format(month_year, "%m")) %>%
    distinct()
head(df_medicare_agg)
```

- Create aggregated version of data for DME users
```{r}
df_dme_agg <- df_dme %>%
    group_by(month_year) %>%
    summarise(value = sum(value),
           month = format(month_year,"%b"),
           year = as.numeric(format(month_year,"%Y")),
           month_num = format(month_year, "%m")) %>%
    distinct()
head(df_dme_agg)
```

## Plot DME users

- Create a step chart for aggregated DME users

```{r}
df_dme_agg %>%
  ggplot(aes(x=month_year, y= value)) + 
  geom_step()+
  scale_x_date(date_breaks = "1 year", date_minor_breaks = "1 month",
             date_labels = "%Y") +
  scale_y_continuous(name = "Number of DME users",
                     breaks = caculate_class(df_dme_agg$value,4),
                     labels = format(caculate_class(df_dme_agg$value,4),
                                     big.mark = ","))+
  # scale_y_reverse()+
  theme_ipsum()+
  theme(
      axis.title.x = element_blank(),
        # axis.title.y = element_blank(),
        panel.grid.minor = element_line(color = "#f0f0f0",size = 0.1),
        panel.grid.major = element_line(color = "#cccccc",size = 0.1))+
  ggtitle("Change of Number of DME users in California") -> dme_stepchart

dme_stepchart
```

- Save the stepchart
```{r}
ggsave(
  here("plot","11_ca_dme_stepchart.png"),
  dme_stepchart,
  # device = cairo_pdf,
  width = 9,
  height = 6,
  scale = 1
)
```

## Combo stepcharts with Medicare and DME users

- Create a new dataframe for both metrics
```{r}
df_dme_agg <- df_dme_agg %>%
    left_join(df_medicare_agg, by = "month_year",
              suffix = c("","_medicare")) %>%
    mutate(pct_dme_medicare = round(value/value_medicare,4)*100) %>%
    select(-value_medicare,-month_medicare, -year_medicare,
           -month_num_medicare)
head(df_dme_agg)
```

- Create a stepchart

```{r}
df_dme_agg %>%
  ggplot(aes(x=month_year, y= pct_dme_medicare)) + 
  geom_step()+
  scale_x_date(date_breaks = "1 year", date_minor_breaks = "1 month",
             date_labels = "%Y") +
  scale_y_continuous(name = "%DME users in Medicare Beneficiaries",
                     breaks = caculate_class(df_dme_agg$pct_dme_medicare,4),
                     labels = format(caculate_class(df_dme_agg$pct_dme_medicare,4),
                                     big.mark = ","))+
  # scale_y_reverse()+
  theme_ipsum()+
  theme(
      axis.title.x = element_blank(),
        # axis.title.y = element_blank(),
        panel.grid.minor = element_line(color = "#f0f0f0",size = 0.1),
        panel.grid.major = element_line(color = "#cccccc",size = 0.1))+
  ggtitle("Change of %DME users in Medicares of California") -> pctdme_stepchart

pctdme_stepchart
```

- Save the stepchart
```{r}
ggsave(
  here("plot","12_ca_dme_pct_stepchart.png"),
  pctdme_stepchart,
  # device = cairo_pdf,
  width = 9,
  height = 6,
  scale = 1
)
```

## DME users in counties deemed high area of WUI

- Define counties deemed high area of WUI
```{r}
high_wui_counties <- read_csv(here("data_private","wui_county_more_than10pct.csv"))
names_high_wui <- high_wui_counties$NAME
```

- Develop aggregation for selected counties
```{r}
df_dme_agg_high_wui <- df_dme %>%
    filter(County %in% names_high_wui) %>%
    group_by(month_year) %>%
    summarise(value = sum(value),
           month = format(month_year,"%b"),
           year = as.numeric(format(month_year,"%Y")),
           month_num = format(month_year, "%m")) %>%
    distinct() %>%
    ungroup() %>%
    mutate(type = "High in WUI")
head(df_dme_agg_high_wui)
```

```{r}
df_dme_agg_not_high_wui <- df_dme %>%
    filter(!(County %in% names_high_wui)) %>%
    group_by(month_year) %>%
    summarise(value = sum(value),
           month = format(month_year,"%b"),
           year = as.numeric(format(month_year,"%Y")),
           month_num = format(month_year, "%m")) %>%
    distinct() %>%
    ungroup() %>%
    mutate(type = "Not High in WUI")
head(df_dme_agg_not_high_wui)
```


- Combine these two dataframes together
```{r}
df_dme_agg_wui_info <- bind_rows(
    df_dme_agg_high_wui,
    df_dme_agg_not_high_wui
)
```

- Calculate Medicare beneficiaries in high WUI counties
```{r}
df_medicare_agg_high_wui <- df_medicare %>%
    filter(County %in% names_high_wui) %>%
    group_by(month_year) %>%
    summarise(value = sum(value),
           month = format(month_year,"%b"),
           year = as.numeric(format(month_year,"%Y")),
           month_num = format(month_year, "%m")) %>%
    distinct() %>%
    ungroup() %>%
    mutate(type = "High in WUI")
head(df_medicare_agg_high_wui)
```


- Calculate Medicare beneficiaries in not high WUI counties
```{r}
df_medicare_agg_not_high_wui <- df_medicare %>%
    filter(!(County %in% names_high_wui)) %>%
    group_by(month_year) %>%
    summarise(value = sum(value),
           month = format(month_year,"%b"),
           year = as.numeric(format(month_year,"%Y")),
           month_num = format(month_year, "%m")) %>%
    distinct() %>%
    ungroup() %>%
    mutate(type = "Not High in WUI")
head(df_medicare_agg_not_high_wui)
```

- Combine these two dataframes
```{r}
df_medicare_agg_wui_info <- bind_rows(
    df_medicare_agg_high_wui,
    df_medicare_agg_not_high_wui
)
```

- Calculate percentage of DME users in the selected counties
```{r}
df_dme_agg_high_wui <- df_dme_agg_high_wui %>%
    left_join(df_medicare_agg_high_wui, by = "month_year",
              suffix = c("","_medicare")) %>%
    mutate(pct_dme_medicare = round(value/value_medicare,4)*100) %>%
    select(-value_medicare,-month_medicare, -year_medicare,
           -month_num_medicare)
head(df_dme_agg_high_wui)
```

- Make a stepchart for `df_dme_agg_wui_info`
```{r}
df_dme_agg_wui_info %>%
  ggplot(aes(x=month_year, y= value, group(type))) + 
  geom_step(aes(color = type))+
  scale_x_date(date_breaks = "1 year", date_minor_breaks = "1 month",
             date_labels = "%Y") +
  scale_y_continuous(name = "DME users in Medicare Beneficiaries",
                     breaks = caculate_class(df_dme_agg_wui_info$value,4),
                     labels = format(caculate_class(df_dme_agg_wui_info$value,4),
                                     big.mark = ","))+
  scale_color_discrete(name = "Counties")+
  theme_ipsum()+
  theme(
      axis.title.x = element_blank(),
        # axis.title.y = element_blank(),
        panel.grid.minor = element_line(color = "#f0f0f0",size = 0.1),
        panel.grid.major = element_line(color = "#cccccc",size = 0.1))+
  ggtitle("Change of DME users in California") -> dme_wui_stepchart

dme_wui_stepchart
```

- Make a stepchart for `df_medicare_agg_wui_info`
```{r}
df_medicare_agg_wui_info %>%
  ggplot(aes(x=month_year, y= value, group(type))) + 
  geom_step(aes(color = type))+
  scale_x_date(date_breaks = "1 year", date_minor_breaks = "1 month",
             date_labels = "%Y") +
  scale_y_continuous(name = "Medicare Beneficiaries",
                     breaks = caculate_class(df_medicare_agg_wui_info$value,4),
                     labels = format(caculate_class(df_medicare_agg_wui_info$value,4),
                                     big.mark = ","))+
  scale_color_discrete(name = "Counties")+
  theme_ipsum()+
  theme(
      axis.title.x = element_blank(),
        # axis.title.y = element_blank(),
        panel.grid.minor = element_line(color = "#f0f0f0",size = 0.1),
        panel.grid.major = element_line(color = "#cccccc",size = 0.1))+
  ggtitle("Change of Medicare Beneficiaries in California") -> medicare_bene_wui_stepchart

medicare_bene_wui_stepchart
```

- Save to PNG files
```{r}
ggsave(
  here("plot","13_ca_dme_wui_compare_stepchart.png"),
  dme_wui_stepchart,
  # device = cairo_pdf,
  width = 6,
  height = 4,
  scale = 1
)
```

```{r}
ggsave(
  here("plot","14_ca_medicare_wui_compare_stepchart.png"),
  medicare_bene_wui_stepchart,
  # device = cairo_pdf,
  width = 6,
  height = 4,
  scale = 1
)
```